\documentclass[a4paper, twoside]{report}

\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{listings}
\usepackage{hyperref}
\hypersetup{colorlinks=false}
\usepackage{lscape}
\usepackage{subfigure}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{eurosym}
\usepackage{float}
%% Sets page size and margins
\usepackage[a4paper,top=3cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

\title{Short-Term Electric Load Forecasting Using Machine Learning \& Neural Networks}
\author{Panagiotis Ntarzanos}
% Update supervisor and other title stuff in title/title.tex

\begin{document}
\input{title/title.tex}
\vspace*{\fill}
\begingroup
\centering
This thesis is dedicated to Elder Athinagoras, for never giving up, on anything. Especially me.
\endgroup
\vspace*{\fill}
\begin{abstract}
\par Load forecasting is an important component for power system's energy management procedure. Since in power systems the next days’ power generation must be scheduled everyday, day-ahead Short-Term Load Forecasting (STLF) is a necessary daily task for power dispatch. Short-Term Load Forecasting has always been one of the most critical, sensitive and accuracy demanding factors of the power systems.  Besides playing a key role in reducing the generation cost, an accurate STLF improves not only the whole system’s economic viability but also its safety, stability and reliability in operation, in order to realize futuristic Smart Power System. 
\par
Electric load forecasting is an important tool, used to ensure that the energy supplied by utilities meets the consumers needs. Load forecasting has always been defined as the science of predicting the future load on a given system, for a specified period of time ahead. Underprediction of STLF leads to insufficient reserve capacity preparation and thus increases the operating cost by using expensive peaking units. On the other hand, overprediction of STLF leads to the unnecessarily large reserve capacity, which is also related to high operating cost. 
\par Since precise load forecasting remains a great challenge, the research work in this master thesis is to  develop efficient and accurate predictive models for electric load demand forecasting in short-term horizon (for 1-2 days ahead), concerning the Greek Electric Network Grid,  by utilizing novel algorithms and up-to-date techniques. More specifically, we developed several models by using Machine Learning techniques and Recurrent Neural Networks, which are considered to be State of the Art algorithms for predictive analytics, not only for normal days but also taking account special days (e.g Holidays, Weekends etc). Validation of our research points out that these techniques generate better results in comparison with the currently used systems predictions. The input data that this thesis uses originate, from IPTO which stands for Independent Power Transmission Operator, and depict the historically hourly actual load of the Greek Electric Network Grid. Moreover, we also used historical load prediction data that Protergia's analysts declared in IPTO's load declaration platform, in order to be able to compare them with the thesis predictions and quantify the improvement.
\\
\\
\\
\\
\\
\\
\textbf{Keywords:} Machine Learning, Deep Learning, Time-series, Short-Term Electricity Load Forecasting, ARIMA, Prophet, LSTM ,Random Forest, Gradient Boosting, LightGBM
\end{abstract}

\renewcommand{\abstractname}{Acknowledgements}
\begin{abstract}
Foremost, I would like to express my sincere gratitude to my supervisor Prof. Panagiotis Louridas for the continuous support of my thesis, for his patience, motivation, enthusiasm and immense knowledge. His guidance helped me in all the time of research and writing of this thesis. \\
\\
I would also like to thank the expert who was involved in the process of this research project Mr. Theodoros Paschos, Protergia's Energy Analyst, who provided me with the appropriate data used in this research. Without his passionate participation and input, this thesis could not have been successfully conducted.\\
\\
Finally, I must express my very profound gratitude to my parents for providing me with unfailing support and continuous encouragement throughout my years of study and through the process of researching and writing this thesis. This accomplishment would not have been possible without them. Thank you.\\
\\
\\
\\
Panagiotis Ntarzanos
\end{abstract}

\tableofcontents
\listoffigures


\input{introduction/introduction.tex}
\input{background/background.tex}
\input{project/project.tex}
\input{evaluation/evaluation.tex}
\input{conclusion/conclusion.tex}
\input{appendix/appendix.tex}

\bibliographystyle{unsrt}
\bibliography{bibs/sample}
\addcontentsline{toc}{chapter}{Bibliography}

\end{document}